\section{Introduction}%
\label{sec:intro}

With the tool UPPAAL Stratego~\cite{David2015UppaalS} it is possible to learn
safe and (near-)optimal strategies for controllers of any system that can be
modelled as a Markov Decision Process (MDP). This has many important
applications, one being the government of cyber-physical systems that often have
important safety requirements to avoid real-world catastrophees while also
needing to perform at a certain level of expectation to satisfy consumers,
citizens or other participants in a network.

For these cases of cyber-physical systems -- especially within critical
infrastructure such as water management, transport or energy distribution ---
another important feature of a controller is that it is explainable and
understandable by a human expert. A well-established technique for representing
machine learning strategies in ways interpretable by humans is with decision
trees, where the path from a root node to a leaf node directly specifies how a
state of the system is evaluated so that a specific action is decided upon.

In UPPAAL Stratego the strategies learned are represented by a collection of
decision trees, one for each action. The leaf nodes of each tree then gives an
evaluation of the expected cost of taking that action in the state under
evaluation. In Q-learning, the Reinforcement Learning technique powering UPPAAL
Stratego, this is called the Q-value and as such, we call the trees Q-trees.
While this approach allows for an online discretization of a continuous state
space and improved approximation of the optimal strategy, it does not retain the
expressive nature of the decision tree representation of its strategy, as each
decision is now an $\argmin$ function over the unique evaluation of the expected
cost of state-action pairs.

In this paper, we therefore suggest a method for converting a set of Q-trees to
a single decision tree representing the same deterministic strategy that the
$\argmin$ over the set of Q-trees would yield. This makes UPPAAL Stratego able
to not only give safe and near-optimal strategies but also to provide them in a
fashion that has the benefits of decision tree representation.

However, we often see that the number of paths in these trees grows very large,
which again makes them exceedingly hard for humans to actually comprehend.
Furthermore, since many controllers of cyber-physical systems will run on
embedded hardware with limited space and memory capacity, large strategies might
be impossible to even run get running in these systems. Therefore, we also
provide an algorithm for minimizing the number of leaf nodes the decision tree
by inspecting the Euclidean state space of the system under consideration and
the partitions that the strategy entails.
